---
title: "TidyTuesday - 2021 Week 29 [Scooby Doo]"
author: "Robert Muwanga"
date: "2025-06-01"
categories: [tidytuesday]
draft: true
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
theme_set(theme_minimal())

# Get dataset
tuesdata <- tidytuesdayR::tt_load(2021, week = 29)
scooby <- tuesdata$scoobydoo
```

## Scooby Doo Episodes

This week's tidytuesday dataset is provided by [plummye](https://www.kaggle.com/williamscooleman) who manually aggregated the Scooby Doo episode data, watching every Scooby-Doo iteration and tracking every variable.

## About the data

The has `r nrow(scooby)` observations across `r ncol(scooby)` variables, running from `r min(scooby$date_aired)` to `r max(scooby$date_aired)`.

```{r}
scooby <- tuesdata$scoobydoo
scooby
```

## Conclusion

Focusing on cleaning one's data set into a tidy dataset helps to set one up for success, making it easier to use functions that comply with the tidy principles in one's analysis. It may take a bit of work to get it right but once its done appropriately, it is quite powerful.
